DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report

### 2022-10-09 Draft-Ped-Metrics-Example.Rmd
adjusts map plot sizing and corrects the points

---
title: "DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report"
output: html_document
fig_caption: yes
bibliography: metrics.bib
date: "`r Sys.Date()`"
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, message = FALSE)
options(scipen=999)

@@

## Purpose

@@

## Introduction

@@

```{r kitsapmap, cache = TRUE}
library(tidyverse)
library(sf)

@@

## Data and Methods

### Data {#datasection}

@@

#### Combining the Data

@@

##### Matching on Road Log Id is good but insufficient

@@

### Method of Evaluation

@@

Because of all of the items that are not available in the road data at this time, this draft makes some assumptions to allow for an evaluation. When the information is known, a more accurate evaluation can be made.

@@

## Results
First, we calculate the score overall for the county and then break it down for other smaller units of area in our community.

### Overall

@@

@@

### Pedestrian Generators

For pedestrian generators, we identify an area that is within 1 mile of the facility and evaluate roads therein in the same manner as the defined county areas, like UGAs and LAMIRDS.

@@

#### Schools
```{r}
pub_schools <- schools %>%
  filter(TYPE == "PUBLIC")

@@

```{r plot_pubschools}
kitsap +
  geom_sf(data = pub_schools2, aes(fill = id), 
          lwd = 0, alpha = 0.3) +
  geom_sf(data = pub_schools, aes(color = 1), color = "black", show.legend = FALSE) +
  theme_void() +
  labs(title = 'Kitsap County Public Schools',
       subtitle = "Schools not labeled with color are not in the unincorporated county",
       caption = "Data source: Kitsap GIS - data in development",
       fill = "Public Schools")

@@

```{r loop_pubschools,include=FALSE,echo=FALSE}
the_schools <- unique(st_set_geometry(pub_schools2, NULL)$id)
loop_df <- st_as_sf(pub_schools2) %>% mutate(tomatch = id)

@@

#### Libraries

@@

```{r plot_libraries}

#lib_circles <- st_intersection(st_buffer(libraries, 5280), st_buffer(county_no_cities,0))

kitsap +
  geom_sf(data = lib2, aes(fill = id), 
          lwd = 0, alpha = 0.3) +
  geom_sf(data = libraries, aes(color = 1), color = "black", show.legend = FALSE) +
  theme_void() +
  labs(title = 'Kitsap County libraries',
       subtitle = "Libraries not labeled with color are not in the unincorporated county",
       caption = "Data source: Kitsap GIS - data in development",
       fill = "Libraries")

@@

#### Transit

@@

```{r plot_transit}
kitsap +
  geom_sf(data = transit_cp2, aes(fill = id), 
          lwd = 0, alpha = 0.3) +
  geom_sf(data = transit_cp, aes(color = 1), color = "black", show.legend = FALSE) +
  theme_void() +
  labs(title = 'Kitsap County Transit Centers and Transfer Points',
       subtitle = "Centers not labeled with color are not in the unincorporated county",
       caption = "Data source: Kitsap Transit Website[@ktransit]",
       fill = "Transit Centers")

@@

```{r, echo=FALSE, results='asis'}
score_breakdowns <- read_csv(paste0(this_dir,"score_breakdowns.csv"))
res <- lapply(c("the_commdists","the_ugas", "the_lamirds", "the_schools", "the_libraries", "the_transits"), function(x) {
  knitr::knit_child(text = c(
    '## `r gsub("_"," ",x)`',
    '',
    '```{r}',
    'getScoreBreakdownForArea(x)',
    '```',
    ''
  ), envir = environment(), quiet = TRUE)
})
cat(unlist(res), sep = '\n')

>>>>

---
title: "DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report"
date: "`r Sys.Date()`"
bibliography: metrics.bib
output: 
  html_document:
    toc: true
    toc_depth: 3
    fig_caption: yes
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center')
options(scipen=999)

@@

# Purpose

@@

# Introduction

@@

```{r kitsapmap, cache = TRUE}

#| fig.align = "center"
#| fig.width = 12
#| fig.height = 8
#| out.width = '100%'

@@

# Data and Methods

## Data {#datasection}

@@

### Combining the Data

@@

#### Matching on Road Log Id is good but insufficient

@@

## Method of Evaluation

@@

Because of all of the items that are not available in the road data at this time, this draft makes some assumptions to allow for an evaluation. When the information is known, a more accurate evaluation can be made.^[Roads with unknown setting - RUCODE is 0, we match them to the Urban road standards. This may be incorrect.]

@@

# Results
First, we calculate the score overall for the county and then break it down for other smaller units of area in our community.

## Overall

@@

Regions are already defined by an area. For pedestrian generators, we identify an area that is within 30 minute walking distance of the facility and evaluate roads therein.^[30 minute walking distance calculated using the mapboxapi function `mb_isochrone`. Isochrones use map data from Mapbox and OpenStreetMap and their data sources. To learn more, visit https://www.mapbox.com/about/maps/ and http://www.openstreetmap.org/copyright.[@mapbox]]

These regions and pedestrian generators are:
 * Commissioner Districts
 * UGAs
 * LAMIRDs
 * Schools
 * Libraries
 * Transit Centers

@@

@@

### Schools
```{r}
pub_schools <- schools %>%
  filter(TYPE == "PUBLIC")

@@

```{r plot_pubschools}
#| fig.width = 8,
#| fig.height = 8,
#| out.width = '100%',
#| dpi = 96

kitsap +
  geom_sf(data = pub_schools2, aes(fill = id), 
          lwd = 0, alpha = 0.3) +
  geom_sf(data = pub_schools, aes(color = 1), color = "black", show.legend = FALSE) +
  theme_void() +
  labs(title = 'Kitsap County Public Schools',
       subtitle = "Schools not labeled with color are not in the unincorporated county",
       caption = "Data source: Kitsap GIS - data in development\n
                  Walking distance polygons from Mapbox",
       fill = "Public Schools")

@@

```{r loop_pubschools,include=FALSE,echo=FALSE, cache=TRUE}
the_schools <- unique(st_set_geometry(pub_schools2, NULL)$id)
loop_df <- st_as_sf(pub_schools2) %>% mutate(tomatch = id)

@@

### Libraries

@@

```{r plot_libraries}
kitsap +
  geom_sf(data = lib2, aes(fill = id), 
          lwd = 0, alpha = 0.3) +
  geom_sf(data = libraries, aes(color = 1), color = "black", show.legend = FALSE) +
  theme_void() +
  labs(title = 'Kitsap County libraries',
       subtitle = "Libraries not labeled with color are not in the unincorporated county",
       caption = "Data source: Kitsap GIS - data in development\n
                  Walking distance polygons from Mapbox",
       fill = "Libraries")

@@

### Transit

@@

```{r plot_transit}
kitsap +
  geom_sf(data = transit_cp2, aes(fill = id), 
          lwd = 0, alpha = 0.3) +
  geom_sf(data = transit_cp, aes(color = 1), color = "black", show.legend = FALSE) +
  theme_void() +
  labs(title = 'Kitsap County Transit Centers and Transfer Points',
       subtitle = "Centers not labeled with color are not in the unincorporated county",
       caption = "Data source: Kitsap Transit Website\n
                  Walking distance polygons from Mapbox",
       fill = "Transit Centers")

@@

```{r, echo=FALSE, results='asis'}
score_breakdowns <- read_csv(paste0(this_dir,"score_breakdowns.csv"))
res <- lapply(c("the_commdists","the_ugas", "the_lamirds", "the_schools", "the_libraries", "the_transits"), function(x) {
  knitr::knit_child(text = c(
    '## `r str_to_title(gsub("_"," ",x))`',
    '',
    '```{r}',
    'getScoreBreakdownForArea(x)',
    '```',
    ''
  ), envir = environment(), quiet = TRUE)
})
cat(unlist(res), sep = '\n')

### 2022-10-09 child_single_geo.Rmd
adjusts map plot sizing and corrects the points

# {{current_geo}}

@@

# need to handle the case where this_sw_sh or this_roads has 0 rows
has_roads <- nrow(this_roads) != 0
has_sw_sh <- nrow(this_sw_sh) != 0
has_both <- has_roads & has_sw_sh
# do not evaluate most of this child if does not
this_comb_total_score <-0
```

@@

if (length(which(eval(as.symbol(the_points)) == this_geo$tomatch)) !=0) {
  # we need to figure out which one
  has_point = TRUE
  tr <- which(eval(as.symbol(the_points)) == this_geo$tomatch)
  if (tr <= length(the_schools)) { # it is a school
    this_point <- st_as_sf(pub_schools) %>%
      filter(NAME == this_geo$tomatch)
  } else if (tr > length(the_schools) & 
             tr <= length(the_schools + length(the_libraries))) { # library
    this_point <- st_as_sf(libraries) %>%
      filter(NAME == this_geo$tomatch)
  } else { #transit
    this_point <- st_as_sf(transit_cp) %>%
      filter(NAME == this_geo$tomatch)
  }
}

@@

```{r, eval=has_both, fig.align = "center", out.width = '100%'}
p1 <- kitsap +
  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.5, show.legend = FALSE) +
  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) +
  geom_sf(data = this_sw_sh, aes(color = as.factor(TYPE)), show.legend = "line") + 
  sw_sh_col +
  theme_void() +
  labs(title = paste("Map for","{{current_geo}}"),
       caption = "Data source: Kitsap GIS - data in development",
       color = "Pedestrian Facilities")

r <- ggplot() +
  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.4, lwd = 0, show.legend = FALSE) +
  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
  geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(color = as.factor(TYPE))) +
  {if(has_point) geom_sf(data = this_point, 
                         aes(color = 1), color = "black", show.legend = FALSE)} +
  sw_sh_col +
  labs(title = "Right",
       color = "Pedestrian Facilities") +
  theme_void()

l <- ggplot() +
  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.4, lwd = 0, show.legend = FALSE) +
  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
  geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(color = as.factor(TYPE))) +
  {if(has_point) geom_sf(data = this_point, 
                         aes(color = 1), color = "black", show.legend = FALSE)} +
  sw_sh_col +
  labs(title = "Left",
       color = "Pedestrian Facilities") +
  theme_void()

@@

(p1 + 
  plot_layout(guides = 'collect') +
  labs(title = paste("Map for","{{current_geo}}"),
       caption = "Data source: Kitsap GIS - data in development",
       color = "Pedestrian Facilities") & 
    theme(legend.position = "bottom") ) +
  (l + theme(legend.position = "none") ) +
  (r + theme(legend.position = "none") )

@@

this_comb_total_score <- sum(this_comb_score2$w_score)/sum(this_comb_score2$the_length)
```

@@

A few segments with a score of 0 are shown here. 
```{r eval=has_both}

@@

```{r}
# add the score to the score breakdowns
if (!exists(deparse(substitute(this_comb_total_score)))){
  this_comb_total_score <- 0
}
score_breakdowns$Score[score_breakdowns$Area == "{{current_geo}}"] <- this_comb_total_score
```

>>>>

#### {{current_geo}}

@@

# need to handle the case where this_sw_sh or this_roads has 0 rows
has_roads <- nrow(this_roads) != 0
has_sw_sh <- nrow(this_sw_sh) != 0
has_both <- has_roads & has_sw_sh
# do not evaluate most of this child if does not

# reset
this_comb_total_score <-0
has0 <- F
has10 <- F
this0 <- ""
this10 <- ""
```

@@

if (length(which(eval(as.symbol(the_points)) == this_geo$tomatch)) !=0) {
  # we need to figure out which one
  has_point = TRUE
  tr <- which(eval(as.symbol(the_points)) == this_geo$tomatch)
  if (tr <= length(the_schools)) { # it is a school
    this_point <- st_as_sf(pub_schools) %>%
      filter(NAME == this_geo$tomatch)
  } else if (tr > length(the_schools) & 
             tr <= length(the_schools) + length(the_libraries)) { # library
    this_point <- st_as_sf(libraries) %>%
      filter(NAME == this_geo$tomatch)
  } else { #transit
    this_point <- st_as_sf(transit_cp) %>%
      filter(NAME == this_geo$tomatch)
  }
}

@@

```{r, eval=has_both}

#| fig.width = 30,
#| fig.height = 8,
#| out.width = '100%',
#| dpi = 96

p1 <- kitsap +
  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.5, show.legend = FALSE) +
  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) +
  geom_sf(data = this_sw_sh, aes(color = as.factor(TYPE)), show.legend = "line") + 
  sw_sh_col +
  theme_void() +
  labs(title = "{{current_geo}}",#paste("Map for","{{current_geo}}"),
       color = "Pedestrian Facilities")

r <- ggplot() +
  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.4, lwd = 0, show.legend = FALSE) +
  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
  geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(color = as.factor(TYPE))) +
  {if(has_point) geom_sf(data = this_point, 
                         aes(color = 1), color = "black", show.legend = FALSE)} +
  sw_sh_col +
  labs(title = "Right",
       color = "Pedestrian Facilities") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

l <- ggplot() +
  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.4, lwd = 0, show.legend = FALSE) +
  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
  geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(color = as.factor(TYPE))) +
  {if(has_point) geom_sf(data = this_point, 
                         aes(color = 1), color = "black", show.legend = FALSE)} +
  sw_sh_col +
  labs(title = "Left",
       color = "Pedestrian Facilities") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

# calculate the number of rows needed to display the legend
this_rows <- max(4,ceiling(n_distinct(as.character(this_sw_sh$TYPE))/3))

@@

((p1 + 
  plot_layout(guides = 'collect') +
  labs(color = "Pedestrian Facilities") & 
    theme(legend.position = "bottom") &
    guides(color=guide_legend(nrow=this_rows,byrow=TRUE))) +
  (l + theme(legend.position = "none") ) +
  (r + theme(legend.position = "none") ) )

@@

this_comb_total_score <- sum(this_comb_score2$w_score)/sum(this_comb_score2$the_length)

# do any with sw sh have score 0?
has0 <- any(this_comb_score$score == 0)
# do any with sw sh have score 10? 
has10 <- any(this_comb_score$score == 10)
this0 <- ifelse(has0, "A few segments where sidewalks or shoulders were matched to a road segment, but the resulting score was 0 are shown here. ","")
this10 <- ifelse(has10, "A few good (score 10) segments are shown here. ","")
```

@@

`r this0`

```{r eval=has_both & has0}

@@

`r this10`

```{r eval=has_both & has10}

roads_10 <- this_comb_score %>%
  filter(score == 10) %>%
  select(setting, road_class, FULL_NAME_R, 
         sidewalk_right, sidewalk_left,
         shoulder_right, shoulder_left,
         standard_sidewalk, standard_shoulder)

knitr::kable(head(roads_10) %>%
               rename(Setting = setting,
                      Class = road_class,
                      Name = FULL_NAME_R,
                      `sidewalk R` = sidewalk_right,
                      `sidewalk L` = sidewalk_left,
                      `shoulder R` = shoulder_right,
                      `shoulder L` = shoulder_left,
                      `standard sidewalk` = standard_sidewalk,
                      `standard shoulder` = standard_shoulder)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```
```{r}
# add the score to the score breakdowns
if (!exists(deparse(substitute(this_comb_total_score)))){
  this_comb_total_score <- 0
}
score_breakdowns$Score[score_breakdowns$Area == "{{current_geo}}"] <- this_comb_total_score
```
